---
title: "Building Distributable Packages"
author: "John Muschelli"
output: 
  ioslides_presentation:
    widescreen: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Outline

- Naming multiple functions
- Package Vignettes
- Rbuildignore
- Including "data" in a package
- Installing from GitHub
- Unit Testing
- Version Numbering
- Checking Packages
- Checking with Travis CI/Appveyor/Coveralls


## Quick naming convention guidelines

These are personal but serve me well:

- All of my functions are lowercase with underscores separating words
- The name of the file is `function_name.R` where the name of the function is `function_name`.  This helps me find things
    - This means I have a **lot** of files in R/ folder.
    - I use the go to file/function search to easily search through them
- If you have multiple functions that should have the same help file, you can use the `@rdname` tag, making sure you still put `@export` at the second function definition.
- Don't make filenames too long (over 25-30 characters max)
- Never put spaces in filenames.


## Example - 2 functions in the same help

See `?log` for example.  
```
#' @title Best title ever
#' @description Best description
#' @export
best = function(){
  print("you're the best around!")
}

#' @rdname best
#' @export
around = function() {
  print("nothing's gonna ever keep you down!")
}
```
## Package Vignettes

Vignettes are package tutorials to show how the package should work

- start with typing `devtools::use_vignette("My_Vignette")`.  Do NOT put spaces in the name.  It's allowable, but just don't.
- Change the title and VignetteEntry tag.
- Write vignette using RMarkdown (making sure to use `library` calls)
    - make sure all packags in `library` calls are in at LEAST Suggests, if not Imports/Depends
- You can build vignettes with `devtools::build_vignettes`
    

## Rbuildignore

The file `.Rbuildignore` has a period in front of it.  That means it's hidden (depending on your settings for folders).  

Go to go to file/function search bar and type `.Rbuildignore`.  Anything in here says "when you go to build R package, don't worry about these files".  Wildcards are fine (like `*`).  Use escapes `\` to escape periods because they are wildcards.

For example, if you have a `README.md` file that is for GitHub but not your package:
```
README\.md$
```
would be in `.Rbuildignore`

## Including data

See http://r-pkgs.had.co.nz/data.html first. From there:

> The most common location for package data is (surprise!) data/. 
> Each file in this directory should be a .RData file created by save() containing a single object (with the same name as the file).

I name mine with `.rda`.  Either use `.rda` or `.RData` - NO OTHER SPELLINGS - it's important.

```{r, eval = FALSE}
blah = matrix(rnorm(100), nrow = 5)
save(blah, file = "data/blah.rda", compression_level = 9)
```

## Documenting data: in `R/blah.R`

```
#' @title Random data, very blah.
#'
#' @description A dataset I made as an illustration.
#'
#' @format A matrix with 20 rows and 5 columns:
#' \describe{
#'   \item{col1}{random col}
#'   \item{col2}{random col}
#'   ...
#' }
"blah"
```

## Documenting "data"

Data here is not always a `data.frame` or `matrix`.  It can be a `lm` or `glm` object.  

You can make sure you reference these specifically by:

```{r, eval = FALSE}
mypkg::blah
```

anywhere in the code.

There are ways to include non-rda files using `inst`, but we won't cover that here


## Version Numbering

From http://r-pkgs.had.co.nz/release.html#release-version use `major.minor.patch.dev`:

* Increment patch, e.g. 0.8.2 for a patch: you’ve fixed bugs without adding any significant new features. 

* Increment minor, e.g. 0.9.0, for a minor release. A minor release can include bug fixes, new features and changes in backward compatibility. 

* Increment major, e.g. 1.0.0, for a major release. This is best reserved for changes that are not backward compatible and that are likely to affect many users. 


## Unit Testing

`use_testthat` - use the `testthat` package to test a package

- if you change function A, does it work with the cases it used to?  There should be a test for that
    
Will give you a `tests` folder and `testthat` folder under it along with a `test.R`

## Example of a test

```r
context("This is the test purpose")

test_that("test_label", {
  expect_success(top(matrix(1:3, nrow = 2)))
})

test_that("test_vectors", {
  expect_error(top(1:3))
})
```

## Checking a Package

Overall, you should check your package to make sure it works.

In increasing order of requirements a package must:

1.  Build - `devtools::build()`
2.  Install - `devtools::install()`
3.  Check - `devtools::check()`
4.  Check with CRAN requirements - `devtools::check(args = '--as-cran')` or `rhub::check_for_cran()`
5.  Check with Bioconductor requirements - `BiocCheck::BiocCheck(package = "package.tar.gz")`

## Checking a Package

When releasing a package to CRAN, it must pass `R CMD check --as-cran`.  This is why we put that option in before.  

- Build -> Check Package - should have no errors, warnings, and likely no notes
- Build -> Build Source Package - this will bundle your package to an installable file.  
- Need to use `R-devel CMD check --as-cran` where `R-devel` is the development version of the packag
    - can do all of this on Travis and Appveyor or Win Builder (https://win-builder.r-project.org/)
- Also see the `rhub` package: https://github.com/r-hub/rhub
    - Makes this easy and all online


## Installing from GitHub

If you have pushed your package to GitHub, you can install it using:

```{r, eval = FALSE}
devtools::install_github("YOURUSERNAME/mypkg")
```
where `mypkg` is the name of the repo.  Make the repo name the same name as the package. It's just easier.  This may be replaced in the future by:
```{r, eval = FALSE}
remotes::install_github("YOURUSERNAME/mypkg")
```

## Checking with Travis CI/Appveyor/Coveralls

`use_travis`, `use_appveyor`, `use_coverage`

These will add YAML files to your packages on Github for testing in Linux/Mac (Travis), Windows (Appveyor), and Code coverage (Coveralls/Travis).

Go to https://travis-ci.org/, sign up with GitHub, sync account, then turn "on" the travis CI for that package.

Go to https://www.appveyor.com/ and do the same thing.

And then https://coveralls.io/ or https://codecov.io/ and do the same thing

PUSH!






